---
title: "Netflix data analysis and visualisation"
output: html_notebook
---

Importing necessary libraries.

```{r}
library(dplyr)
library(ggplot2)
library(scales)
```

Importing (reading) the data of Best movies and Shows. Grabbing high-rated movies/shows (min score 6.9)

```{r}
dataBestMovies <- read.csv("C:/Users/Asus/OneDrive/Desktop/Group project - Netflix/Best Movies Netflix.csv")
dataBestShows <- read.csv("C:/Users/Asus/OneDrive/Desktop/Group project - Netflix/Best Shows Netflix.csv")
```

Exploring data & checking for missing values.
Get data overview & identify missing values.
Printing summary statistics and checking for missing data.
Analyzing data distribution and identifying missing values.

```{r}
has_missingMovie <- sum(is.na(dataBestMovies)) > 0
has_missingShow <- sum(is.na(dataBestShows)) > 0
dataBestMovies$MAIN_GENRE <- factor(dataBestMovies$MAIN_GENRE)
dataBestMovies$MAIN_PRODUCTION <- factor(dataBestMovies$MAIN_PRODUCTION)
dataBestShows$MAIN_GENRE <- factor(dataBestShows$MAIN_GENRE)
dataBestShows$MAIN_PRODUCTION <- factor(dataBestShows$MAIN_PRODUCTION)
print(summary(dataBestMovies))
print(summary(dataBestShows))
```
This code block converts the release dates of movies and shows into their corresponding decades. This will likely be used for further analysis where grouping by decade is important.

```{r}
group_by_date <- function(DOF) {
  start_decade <- seq(from = 1980, to = 2020, by = 10)
  end_decade <- start_decade + 9
  for (i in length(start_decade):1) {
    if (DOF >= start_decade[i] & DOF < end_decade[i]) {
      decade <- paste0(start_decade[i], 's')
      return (decade)
    }
  }
  if (DOF < start_decade[1]) {
    return("Released before 1980")
  } else {
    return ("Released after 2020")
  }
}
```

Implement DECADES column derived from *group_by_date* function.

```{r}
dataBestMovies$DECADES <- sapply(dataBestMovies$RELEASE_YEAR, group_by_date)
dataBestShows$DECADES <- sapply(dataBestShows$RELEASE_YEAR, group_by_date)
```

Calculating the average score, number of movies for each decade and main genre combination.

```{r}
grouppedMovies <- dataBestMovies %>%
  group_by(MAIN_GENRE, DECADES) %>%
  summarise(SCORE = mean(SCORE),
            COUNT_OF_MOVIES = n())

grouppedShows <- dataBestShows %>%
  group_by(MAIN_GENRE, DECADES) %>%
  summarise(SCORE = mean(SCORE),
            COUNT_OF_SHOWS = n())
```
Visualization of Movies

Popularity of movie genres across different decades using pie charts. Each pie chart will represent a specific decade, with slices depicting the proportion of movies belonging to various genres. This visualization will highlight the dominant genres favored by producers throughout time, with a particular focus on Drama's position as the "absolute champion."

```{r}
grouppedMovies$DECADES <- factor(grouppedMovies$DECADES, levels = c("Released before 1980", "1980s", "1990s", "2000s", "2010s", "2020s", "Released after 2020"))

ggplot(data = grouppedMovies, aes(x = "", y = COUNT_OF_MOVIES, fill = MAIN_GENRE)) +
  geom_col(width = 1, position = 'fill') +
  facet_wrap(.~DECADES, nrow=3, ncol = 3) +
  labs(title = "Movie genres by their decades",
       x = "Main Genre",
       y = "Count") +
  scale_x_discrete() +
  coord_polar(theta = "y", start = 0) +
  theme_minimal() +
  theme(strip.text.x = element_text(size = 5))
```
Science fiction, westerns, horror, and crime emerge as the top four most popular genres in your analysis. This variety suggests a broad audience interest across genres that explore imaginative worlds, historical settings, thrilling suspense, and complex characters.The variable that is chosen for this analysis is the number of votes.

```{r}
voters_favourite_Ms <- dataBestMovies %>%
  group_by(MAIN_GENRE) %>%
  summarise(AVERAGE_VOTERS = mean(NUMBER_OF_VOTES),
            SCORE = max(SCORE)) %>%
  arrange(desc(AVERAGE_VOTERS)) %>%
  head(4)

print(voters_favourite_Ms)

ggplot(data = voters_favourite_Ms, aes(x = reorder(MAIN_GENRE, -AVERAGE_VOTERS), y = AVERAGE_VOTERS/1000)) +
  geom_bar(fill = "pink", stat = "identity") +
  labs(title = "Popularity of genres among viewers (Movies)",
       x = "MAIN GENRE", y = "NUMBER OF VOTERS") +
  coord_cartesian(ylim = c(150, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = voters_favourite_Ms, aes(x = reorder(MAIN_GENRE, -SCORE),
                                        y = SCORE)) +
  geom_bar(fill = "hotpink", stat = "identity") +
  labs(title = "Popularity of genres among viewers (TV Shows)",
       x = "MAIN GENRE", y = "SCORE") +
  coord_cartesian(ylim = c(7.5, 10)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Visualization of TV Shows

Genres with consistently high average scores across decades might be considered universally popular. Genres with fluctuating average scores across decades might indicate evolving audience tastes. Decades with a specific genre consistently scoring high might suggest a period of dominance for that genre.

```{r}
grouppedShows$DECADES <- factor(grouppedShows$DECADES, levels = c("Released before 1980", "1980s", "1990s", "2000s", "2010s", "2020s", "Released after 2020"))

ggplot(data = grouppedShows, aes(x = "", y = COUNT_OF_SHOWS, fill = MAIN_GENRE)) +
  geom_col(width = 1, position = "fill") +
  facet_wrap(.~DECADES, nrow = 3, ncol = 3) +
  labs(title = "Number of TV Shows by Genre per Decade",
       x = "Main Genre",
       y = "Count") +
  scale_x_discrete() +
  coord_polar("y", start = 0) + 
  theme_minimal() +
  theme(strip.text.x = element_text(size = 5))
```
Science fiction, action, thriller, and drama emerge as the top four most popular genres in your analysis. This variety suggests a broad audience interest across genres that explore imaginative worlds, historical settings, thrilling suspense, and complex characters.The variable that is chosen for this analysis is the number of votes.

```{r}
voters_favourite_Shs <- dataBestShows %>%
  group_by(MAIN_GENRE) %>%
  summarise(AVERAGE_VOTERS = mean(NUMBER_OF_VOTES),
            SCORE = max(SCORE)) %>%
  arrange(desc(AVERAGE_VOTERS)) %>%
  head(4)

print(voters_favourite_Shs)

ggplot(data = voters_favourite_Shs, aes(x = reorder(MAIN_GENRE, -AVERAGE_VOTERS), y = AVERAGE_VOTERS/1000)) +
  geom_bar(fill = "purple", stat = "identity") +
  labs(title = "Popularity of genres among viewers (TV Shows)",
       x = "MAIN GENRE", y = "NUMBER OF VOTERS") +
  coord_cartesian(ylim = c(100, 150)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = voters_favourite_Shs, aes(x = reorder(MAIN_GENRE, -SCORE),
                                        y = SCORE)) +
  geom_bar(fill = "plum", stat = "identity") +
  labs(title = "Popularity of genres among viewers (TV Shows)",
       x = "MAIN GENRE", y = "SCORE") +
  coord_cartesian(ylim = c(7.5, 10)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```




